---
title: "Rmarkdown Lab Assignment"
author: "Avetis Mishegyan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

# Introduction

Today we are going to practice *RMarkdown* in lab based on the following instructions:  

1. Take any of the data sheets you have already worked with in class and create **at least**  
 **one table and one figure in an Rmarkdown file.**  
2. Create the most beautiful RMarkdown output you can during lab hours in either an  
 html, github doc, or pdf format.  
3. Make sure you have headings with clear explanations of what you are doing.  
4. Practice using markdown formatting with bold, italics, and lists (use each of these at  
 least once).  
5. Make sure your outputs and scripts are saved in the appropriate folders. Keep proper  
 coding etiquette (figures need to be pub quality as usual and you need to comment  
 your code).  

## Load Libraries

Here is where we will load any libraries that we will be using. These libraries contain use full packages that are collections of R functions, data, and compiled code in a well-defined format.

```{r}
library(tidyverse)
library(here)
library(viridis)
library(kableExtra)
```

## Load Data

Here is where we will load (ie. pull) our data as a ".csv" format into our R script. Additionally, this is where we would also inspect our data.

```{r}
topt_data <- read_csv(here("week_6", "data", "Topt_data.csv")) # loads the data from a .csv file
glimpse(topt_data) # allows us to inspect the data

site_data <- read_csv(here("week_6", "data", "site.characteristics.data.csv"))
glimpse(site_data)
```

## Data Analysis

Here is where we will start our data analysis whether that be calculations, graphs, etc.

```{r lab_practice_RMarkdown_2023_03_02, fig.width=10, fig.height=5, fig.align='center'}
topt_site_data <- left_join(topt_data, site_data) %>% # keeps all data from topt_data and matches data from site_data
  drop_na() %>% # drops all NA's from data
  mutate(rate_type = case_when(rate.type == "C" ~ "Calcification", # renames rate type to full name
                               rate.type == "R" ~ "Respiration",
                               rate.type == "GP" ~ "Gross Photosynthesis"))

glimpse(topt_site_data) # let's check out our mutation or change below:

# Here is a more beautiful visualization below:
topt_site_data %>%
  head() %>% # allows us to view the first 6 rows
  kbl() %>% # makes a kable table
  kable_classic() %>% # sets kable table theme to classic
  column_spec(13, bold = T, color = "white", background = "red") # highlights, changes font color, and bolds font of column 13

topt_site_graph <- topt_site_data %>% 
  ggplot(mapping = aes(x = name,
                       y = rate_type,
                       fill = Topt)) +
  geom_tile() + # heat map function
  labs(x = "Site Name",
       y = "Rate Type",
       fill = "Thermal Optima (°C)",
       title = "Thermal Optima by Rate Type and Loaction") +
  scale_fill_viridis(discrete = FALSE) + # sets heat map colors
  theme_classic() + # sets theme to classic
  theme(plot.title = element_text(hjust = .5)) # centers plot title

topt_site_graph # allows us to view our graph
```
