---
title: "Iterative Coding Lab Assignment"
author: "Avetis Mishegyan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

## Introduction

Today we are going to practice iterative coding on temperature and light data to:

* Bring in all 4 files
* Calculate the mean and standard deviation of both temperature and light for each tide pool
* Use both a for loop and map() functions in your script

## Load Libraries

Here is where we will load any libraries that we will be using. These libraries contain use full packages that are collections of R functions, data, and compiled code in a well-defined format.

```{r}
library(tidyverse)
library(here)
```

## Load Data & Data Processing

Here is where we will load (ie. pull) our data as a ".csv" format into our R script. Additionally, this is where we would also inspect our data. In addition, here is were we will process the data since we are using loops.

### Listing Files in a Directory & Pre-Allocating Space for the Loop

```{r}
# List Files in a Directory
hw_path <- here("week_14", "data", "homework") # points to the location on the computer of the folder
hw_files <- dir(path = hw_path, pattern = ".csv") # lists all the files in that path with a specific pattern (ie. ending in .csv)
hw_files # prints list

# Pre-Allocating Space for the Loop
hw_data <- data.frame(matrix(nrow = length(hw_files), ncol = 5)) # makes an empty dataframe that has one row for each file and 3 columns
colnames(hw_data) <- c("filename", "mean_temp", "sd_temp", "mean_light", "sd_light") # names dataframe columns
hw_data # prints table
```

### Making For Loop

```{r}
for (i in 1:length(hw_files)){ # loops over 1:4 the number of files
  raw_hw_data <- read_csv(paste0(here(hw_path, hw_files[i]))) # reads in data
  # glimpse(raw_hw_data) # allows us to inspect the data
  hw_data$filename[i] <- hw_files[i] # adds file names to filename column
  hw_data$mean_temp[i] <- mean(raw_hw_data$Temp.C, na.rm = TRUE) # calcs mean temp of each file
  hw_data$mean_light[i] <- mean(raw_hw_data$Intensity.lux, na.rm = TRUE)
  hw_data$sd_temp[i] <- sd(raw_hw_data$Temp.C, na.rm = TRUE) # calcs standard deviation temp of each file
  hw_data$sd_light[i] <- sd(raw_hw_data$Intensity.lux, na.rm = TRUE)
}

hw_data
```

### Accomplishing For Loops but with Map Functions

#### Bringing in Files using Purrr
```{r}
hw_files_purrr <- dir(path = hw_path, pattern = ".csv", full.names = TRUE) # save the entire path name
hw_files_purrr
```

#### Reading in the Files & Calculations

```{r}
hw_data_purrr <- hw_files_purrr %>% 
  set_names() %>% # set's the id of each list to the file name
  map_df(read_csv, .id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>% # groups by filename column
  summarise(mean_temp = mean(Temp.C, na.rm = TRUE), # calcs mean temp of each file
            sd_temp = sd(Temp.C, na.rm = TRUE), # calcs standard deviation temp of each file
            mean_light = mean(Intensity.lux, na.rm = TRUE),
            sd_light = sd(Intensity.lux, na.rm = TRUE))

hw_data_purrr
```

