---
title: "Functional Programing Lab"
author: "Avetis Mishegyan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

## Introduction

Today we are going to make a use any data set of our choosing to make two useful functions:

* One that is a plot
* One that is not
* Turn in your functions using RMarkdown with at least 2 examples for each function

## Load Libraries

Here is where we will load any libraries that we will be using. These libraries contain use full packages that are collections of R functions, data, and compiled code in a well-defined format.

```{r}
library(tidyverse)
library(here)
library(alphonse) # for alphonse color palette
```

## Load Data

Here is where we will load (ie. pull) our data as a ".csv" format into our R script. Additionally, this is where we would also inspect our data.

```{r}
undil_chloro_data <- read_csv(here("week_11", "data", "undiluted_turmeric_flood_chlorophyll_test_2022_12_12.csv")) # loads the data from a .csv file
glimpse(undil_chloro_data) # allows us to inspect the data

dil_chloro_data <- read_csv(here("week_11", "data", "diluted_turmeric_flood_chlorophyll_test_2022_12_12.csv"))
glimpse(dil_chloro_data)
```

## Data Analysis

### Total Chlorophyll Function

```{r}
total_chlorophyll <- function(abs_645, abs_663) { # name of function, and list of inputs (ie. arguments)
  chloro_a <- 12.7*abs_663 - 2.69*abs_645 # code you have developed in body of the function
  chloro_b <- 22.9*abs_645 - 4.68*abs_663
  total <- chloro_a + chloro_b
  return(total) # returns values we want
}

total_chlorophyll(.2, .8) # test of function
```

### Total Chlorophyll Calculations for Undiluted Samples

```{r}
total_calc_undil <- undil_chloro_data %>%
  pivot_wider(names_from = abs, # pivots the data wider
              values_from = chloro_undil) %>%
  mutate(total_undil = total_chlorophyll(abs_645, abs_663)*5) %>% # calculates total chlorophyll into new column
  group_by(treatment) %>% # groups data by values in the treatment column
  summarise(mean_total_undil = mean(total_undil, na.rm = TRUE), # computes a table of summarized data
            sd_total_undil = sd(total_undil, na.rm = TRUE))
glimpse(total_calc_undil)
```


### Total Chlorophyll Calculations for Diluted Samples

```{r}
total_calc_dil <- dil_chloro_data %>%
  pivot_wider(names_from = abs,
              values_from = chloro_dil) %>%
  mutate(total_dil = total_chlorophyll(abs_645, abs_663)*5*10) %>%
  group_by(treatment) %>%
  summarise(mean_total_dil = mean(total_dil, na.rm = TRUE),
            sd_total_dil = sd(total_dil, na.rm = TRUE))
glimpse(total_calc_dil)
```

### Bar Plot Function

```{r}
bar_plot <- function(data, x, y, fill, sd, lab_x, lab_y, title){ 
ggplot(data, aes(x = {{x}},
                 y = {{y}}, 
                 fill = {{fill}})) +
  geom_bar(stat = "identity") + # bar plot function
  geom_errorbar(aes(ymin = {{y}} - {{sd}}, # error bar function
                    ymax = {{y}} + {{sd}}),
                width = .2) + # adjusts width of error bars
  labs(x = lab_x,
       y = lab_y,
       title = title) +
  scale_fill_manual(values = alphonse("demon")) + # colors bars using alphonse color palette
  theme_classic() + # sets plot theme to classic
  theme(legend.position = "none") + # removes legend
  theme(plot.title = element_text(hjust = .5)) # centers plot title
}
```

### Total Chlorophyll Bar Plot for Undiluted Samples

```{r total_undiluted_bar, fig.width=10, fig.height=5, fig.align='center',  fig.cap="Figure 1. The mean total chlorophyll of undiluted extractions of all biological replicates consisting of 3 separate leaf samples for each biological replicate for each treatment is represented above. Biological replicates consisted of control: n = 3 and flood: n = 3. Error bars above represent standard deviation."}
bar_plot(data = total_calc_undil,
         x = treatment,
         y = mean_total_undil,
         fill = treatment,
         sd = sd_total_undil,
         lab_x = "Treatment",
         lab_y = "Mean Total Chlorophyll (mg)",
         title = "Turmeric Mean Total Chlorophyll Undiluted Flood v. Control")
```

### Total Chlorophyll Bar Plot for Diluted Samples

```{r total_diluted_bar, fig.width=10, fig.height=5, fig.align='center',  fig.cap="Figure 2. The mean total chlorophyll of diluted extractions of all biological replicates consisting of 3 separate leaf samples for each biological replicate for each treatment is represented above. Biological replicates consisted of control: n = 3 and flood: n = 3. Error bars above represent standard deviation."}
bar_plot(data = total_calc_dil,
         x = treatment,
         y = mean_total_dil,
         fill = treatment,
         sd = sd_total_dil,
         lab_x = "Treatment",
         lab_y = "Mean Total Chlorophyll (mg)",
         title = "Turmeric Mean Total Chlorophyll Diluted Flood v. Control")
```