---
title: "Good Plot vs. Bad Plot Lab Assignment"
author: "Avetis Mishegyan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

## Introduction

Today we are going to make a good plot and bad plot for lab as a competition.  

**Requirements**:

* Create 2 versions of similar plots using any data set
* Plots must be made using ggplot2
  * post-processing with Adobe Illustrator or similar is not allowed
* Submit your homework by posting a markdown document containing
  * data
  * plots
  * code you used to create them
  * discussion of the plot within the markdown doc to github
* Bad plot should be as bad as possible based on the readings from Claus Wilke’s (Ch.1-5) and Healy's (Ch.1)
textbook describe the many ways that a plot can become:
  * misleading
  * difficult to interpret
  * just plain ugly
  * you should use all of the tools that you can
  * explain in text all of the ways in which the bad plot is bad (bullet points are fine, but clearly articulate each principle of data visualization that you have broken)
* Good plot should be based on the same or similar data to the first plot, but it should be good:
  * write briefly about why this is a successful plot

## Load Libraries

Here is where we will load any libraries that we will be using. These libraries contain use full packages that are collections of R functions, data, and compiled code in a well-defined format.

```{r}
library(tidyverse)
library(here)
library(tidytuesdayR)
library(ggplot2)
library(packcircles) # for packed circle plots
library(ggpubr) # package for making beautiful ggplot's
library(jpeg) # for reading in jpeg images
library(forcats) # for reordering within columns
```

## Load Data

Here is where we will load (ie. pull) our data from GitHub into our R script. Additionally, this is where we would also inspect our data.

```{r}
tuesdata <- tidytuesdayR::tt_load('2019-01-29') # reads in data using the tidytuesdayR package

clean_cheese <- tuesdata$clean_cheese
glimpse(clean_cheese) # glimpse allows us to inspect the data
```

## Data Analysis

### Data Manipulation for Bad Plot

```{r}
bad_cheese <- clean_cheese %>%
  select(ChEdDaR = Cheddar, # selects and or renames columns
         `AmErIcAn OtHeR` = `American Other`,
         MoZzArElLa = Mozzarella,
         `ItAlIaN OtHeR` = `Italian other`,
         SwIsS = Swiss,
         BrIcK = Brick,
         MuEnStEr = Muenster,
         `CrEaM aNd NeUfChAtEl` = `Cream and Neufchatel`,
         BlEe = Blue,
         `OtHeR DaIrY ChEeSe` = `Other Dairy Cheese`,
         `PrOcEsSeD ChEeSe` = `Processed Cheese`,
         `FoOdS aNd SpReAdS` = `Foods and spreads`) %>%
  summarise(ChEdDaR = mean(ChEdDaR, na.rm = TRUE), # calculates the mean and excludes any NA's
            `AmErIcAn OtHeR` = mean(`AmErIcAn OtHeR`, na.rm = TRUE),
            MoZzArElLa = mean(MoZzArElLa, na.rm = TRUE),
            `ItAlIaN OtHeR` = mean(`ItAlIaN OtHeR`, na.rm = TRUE),
            SwIsS = mean(SwIsS, na.rm = TRUE),
            BrIcK = mean(BrIcK, na.rm = TRUE),
            MuEnStEr = mean(MuEnStEr, na.rm = TRUE),
            `CrEaM aNd NeUfChAtEl` = mean(`CrEaM aNd NeUfChAtEl`, na.rm = TRUE),
            BlEe = mean(BlEe, na.rm = TRUE),
            `OtHeR DaIrY ChEeSe` = mean(`OtHeR DaIrY ChEeSe`, na.rm = TRUE),
            `PrOcEsSeD ChEeSe` = mean(`PrOcEsSeD ChEeSe`, na.rm = TRUE),
            `FoOdS aNd SpReAdS` = mean(`FoOdS aNd SpReAdS`, na.rm = TRUE),) %>%
  pivot_longer(cols = ChEdDaR:`FoOdS aNd SpReAdS`, # pivots data into new columns as long format data
               names_to = "type",
               values_to = "mean")
glimpse(bad_cheese)
```

### Data Manipulation for Good Plot

```{r}
good_cheese <- clean_cheese %>%
  select(Cheddar,
         `American Other`,
         Mozzarella,
         `Italian Other` = `Italian other`,
         Swiss,
         Brick,
         Muenster,
         `Cream and Neufchatel`,
         Blue,
         `Other Dairy Cheese`,
         `Processed Cheese`,
         `Foods and Spreads` = `Foods and spreads`) %>%
  pivot_longer(cols = Cheddar:`Foods and Spreads`,
               names_to = "type",
               values_to = "values") %>%
  mutate(type = fct_relevel(type, c("Cheddar", # reorders type column the way listed below
                                    "American Other",
                                    "Mozzarella",
                                    "Italian Other",
                                    "Swiss",
                                    "Brick",
                                    "Muenster",
                                    "Cream and Neufchatel",
                                    "Blue",
                                    "Other Dairy Cheese",
                                    "Processed Cheese",
                                    "Foods and Spreads"))) # don't need '' once columns are pivoted into a single column, can check names by typing unique(good_cheese$type)
glimpse(good_cheese)
```

## Plotting Data

### Bad Plot

```{r bad_plot, fig.width=20, fig.height=10, fig.align='center', fig.cap="FiGuRe 1. BaD pLoT oF cLeAn ChEeSe DaTa."}
imgfile <- readJPEG(here("week_10", "data", "cheese.jpeg")) # reads in photo

data <- bad_cheese # creates data frame

packing <- circleProgressiveLayout(data$mean,
                                   sizetype = 'area') # generates the layout, and gives its center (x and y) and radius (proportional of the value)

data <- cbind(data, packing) # adds packing info (line 174) to data frame (line 173)

dat.gg <- circleLayoutVertices(packing, npoints = 50) # finds/sets center and radius to coordinates of a circle

bad_plot <- ggplot() + # makes the plot
  background_image(imgfile) + # adds image to background of ggplot
  geom_polygon(data = dat.gg, aes(x = x, # makes the bubbles
                                  y = y,
                                  group = id,
                                  fill = as.factor(id),
                                  color = "black",
                                  alpha = .6)) +
  geom_text(data = data, aes(x = x, # adds text in center, and adjusts its size
                             y = y,
                             size = radius, # should size by mean, but looks worse with radius
                             label = type)) +
  labs(title = "AmErIcAn ChEeSe CoNsUmPtIoN (LbS/PeRsOn) 1970-2017") +
  scale_fill_grey() + # sets scale fill color to grey
  theme_void() + # sets theme to void
  theme(plot.title = element_text(hjust = .5)) + # centers plot title
  coord_fixed() # fixes aspect ratio to make ovals into circles
bad_plot # prints the plot
```

This is a bad plot due to:

* Bad Taste:
  * tasteless circular packing graph
  * unnecessary legends
  * ugly mocking Spongebob Squarepants text
  * chart junk = unnecessary plot background
* Bad Data:
  * uninformative/ misleading title
  * uninformative figure caption
  * no way to measure samples (no axis scales)
* Bad Perception:
  * differences in sizes of circles are hard to read/ compare
  * circles not aligned
  * poor fill color choice (gray's = no pop/ contrast)
  * poor choice of background = text blends in

### Good Plot

```{r good_plot, fig.width=10, fig.height=5, fig.align='center',  fig.cap="Figure 2. This figure represents the median American consumption of 12 cheese types in total from 1970 to 2017. This data comes from Tidy Tuesday of week 5 on 2019-01-29 sourced from the USDA."}
good_plot <- good_cheese %>%
  ggplot(aes(x = type, # plot function
             y = values,
             fill = type)) +
  geom_boxplot() + # box plot function
  labs(x = "Cheese Type",
       y = "Cheese Consumption (lbs/person)",
       title = "American Consumption of Cheeses from 1970-2017") +
  theme_classic() + # sets theme to classic
  theme(plot.title = element_text(hjust = .5, # centers plot title
                                  size = 15), # sizes plot tile
        axis.text.x = element_text(angle = 45, # angles x-axis labels
                                   vjust = .8, # moves x-axis labels horizontally
                                   hjust = .8), # moves x-axis labels vertically
        legend.position = "none") # removes legend
good_plot
```

This is a good plot due to:

* appropriate plot type for the data
* simple graph:
  * removed grid lines
  * removed unnecessary legends
  * removed unnecessary background
* easy to read text
* informative title and figure caption
* labeled axes
* can measure changes in samples
* color helps distinguish/ visually pops cheese types
* can measure differences between cheese types