---
title: "Maps Lab Assignment"
author: "Avetis Mishegyan"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

## Introduction

Today we are going to make a map using RMarkdown with one of the following datasets from past Tidy Tuesdays:  
1. [Plastic pollution](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-01-26)  
2. [Cats on the move!](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-31)  
3. [Where is Bigfoot!](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-09-13)  
4. [Meteorites](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-06-11)  

The following script below will use the Bigfoot dataset. Additionally, we are making sure, as usual, that our scripts and output are saved appropriately.

## Load Libraries

Here is where we will load any libraries that we will be using. These libraries contain use full packages that are collections of R functions, data, and compiled code in a well-defined format.

```{r}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapdata)
library(mapproj)
```

## Load Data

Here is where we will load (ie. pull) our data as a ".csv" format into our R script. Additionally, this is where we would also inspect our data.

```{r, results="hide"}
tuesdata <- tidytuesdayR::tt_load('2022-09-13') # reads in data using the tidytuesdayR package
```
```{r}
bigfoot <- tuesdata$bigfoot

head(bigfoot) # head shows only first 6 row of the data
glimpse(bigfoot) # glimpse allows us to inspect the data
```

## Data Analysis

Here is where we will start our data analysis whether that be calculations, graphs, etc.

### Data Manipulation - Filter by California

```{r}
cali_bigfoot <- bigfoot %>%
  drop_na() %>% # drops all NA's
  filter(state == "California") # filters out data from California only
glimpse(cali_bigfoot) # let's check out our mutation or change below:

counties <- map_data("county") # gets data for counties in the USA
glimpse(counties)
```

### Plot Data - Bigfoot Sightings

```{r lab_assignment_maps_2023_03_07, fig.width=20, fig.height=10, fig.align='center'}
counties %>% # uses the counties dataset
  filter(region == "california") %>% # filters out just the California data
  ggplot() +
  geom_polygon(aes(x = long, # function that makes a map of county data
                   y = lat,
                   group = group),
               fill = "green",
               color = "black") +
  geom_point(data = cali_bigfoot, # function that adds a point at Bigfoot sightings
             aes(x =longitude,
                 y = latitude,
                 size = number)) + # sets size of points on the map by number of sightings
  labs(size = "# Bigfoot Sightings:") + # sets legend name
  coord_map() + # sets map projection to account for 2D visualization of a globe
  theme_void() + # sets graph theme to void
  scale_fill_gradient(trans = "log10") # changes axes to log scale for easier interpretation
```