---
title: "Working with Factors Lab Assignment"
author: "Avetis Mishegyan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

## Introduction

Today we are going to clean up data using skills from working with words and factors to:

* Clean up issues with the character data
* Reorder factors into orders that make sense
* Make column headers better
* Make a plot with one of the axes as a factor

## Load Libraries

Here is where we will load any libraries that we will be using. These libraries contain use full packages that are collections of R functions, data, and compiled code in a well-defined format.

```{r}
library(tidyverse)
library(here)
library(janitor) # for cleaning up data
library(viridis) # for viridis color palette
```

## Load Data

Here is where we will load (ie. pull) our data as a ".csv" format into our R script. Additionally, this is where we would also inspect our data.

```{r}
intertidal_data <- read_csv(here("week_13", "data", "intertidaldata.csv")) # loads the data from a .csv file
glimpse(intertidal_data) # allows us to inspect the data
```

## Data Analysis

### Cleaning Data
```{r}
clean_intertidal_data <- intertidal_data %>% 
  clean_names() %>% # cleans names of columns
  mutate(quadrat = str_replace_all(quadrat, "[^a-zA-Z]", ""), # removes non-characters that aren't btw a-z
         quadrat = str_trim(quadrat), # removes all white space
         quadrat = factor(quadrat, levels = c("Low", "Mid", "High"))) # converts to a factor and reorders
glimpse(clean_intertidal_data)
```

### Plotting Data

```{r clean_intertidal_plot, fig.width=5, fig.height=5, fig.align='center', fig.cap="Figure 1. This figure represents the percent coverage of algae in each ocean tide quadrat by location."}
clean_intertidal_plot <- clean_intertidal_data %>% 
  drop_na() %>% # removes all NA's
  ggplot(aes(x = quadrat,
             y = site,
             fill = algae)) +
  geom_tile() + # heat map function
  labs(x = "Ocean Tide Quadrat",
       y = "Location",
       fill = "% Cover in Quadrat",
       title = "Algae Population") +
  scale_fill_viridis(discrete = FALSE) + # sets heat map colors
  theme_classic() + # sets theme to classic
  theme(plot.title = element_text(hjust = .5)) # centers plot title
clean_intertidal_plot # prints plot
```